A Dearth of Young Stars and Planets in the Kepler Field

### 2024-05-27 paper/ms.tex
update ms

Stars that formed in the same birth cluster are the gold standard for
the astronomical age scale \citep{Soderblom_2010}.  Before Gaia, the
few open clusters known in the Kepler field had been
visually identified by \citet{1864RSPT..154....1H}.  Gaia proper
motions and distances have enabled the discovery of stellar ensembles
that are more diffuse, but which nonetheless share a common age based
on e.g.~isochrone, rotation, and lithium dating
\cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.

@@

The point of Figure~\ref{fig:agescalecompone} is hopefully clear:
until this study, the ages of the youngest Kepler stars and planets were not 
well-measured.
While previous studies reported accurate ages for
$\gtrsim$1\,Gyr stars, ages of sub-gigayear stars have historically been
overestimated by 0.3--2\,dex, with severely understated uncertainties.  
For isochrone and kinematic ages, this is because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages in
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is
caused by details of the calibration methodologies.  The
\texttt{kiauhoku} spin-down model for instance has known
``stretches'' and ``compressions'' in its age scale relative to observed open clusters
\citep[see][Sec. 7.3]{2023ApJ...952..131M}.
The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} appears more accurate than the former
two models, because it 
was directly fit to reproduce the cluster sequences known at the time.  However the
uncertainties from all of these methods appear to be underestimated.  This
is likely because
they do not marginalize over
the range of rotation periods that stars of a fixed
mass and age might have.

@@

The age distributions of Kepler stars and KOI hosts in Figure~\ref{fig:hist_tgyro} appear
similar.  One diagnostic for whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
range of $\log_{\rm 10}p$ (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentile)
spanned -3.9 to -1.2 for the K-dwarfs only,
and -5.2 to -2.2 for all Kepler stars.  This degree of similarity suggests 
that although weak differences may be present,
the ``common wisdom'' that young transiting planets are hard to detect
because their host stars are photometrically variable requires some
nuance.  If true, this statement only holds for the very youngest
stars, where we see only a marginal underabundance in the KOI host age
distribution relative to that of the overall stellar sample.
At the age resolution of this study, 
the selection bias against sub-gigayear planets seems small.

@@

We can quantify the relative counts of stars as a function of age by
labelling stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr as ``young'',
``intermediate-age'', and ``old''.  A simple counting exercise from
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.  Focusing only on the tails of the
distributions (0-0.3\,Gyr and 2.7-3\,Gyr), the implication is that the
formation rate of stars in the Kepler field decreased by a factor of
\ratiosfr$\pm$\uncratiosfr\ over the past three billion years.

>>>>

Stars that formed in the same birth cluster are the gold standard for
the astronomical age scale \citep{Soderblom_2010}.  Before Gaia, the
few open clusters known in the Kepler field had been
visually identified by \citet{1864RSPT..154....1H}.  Gaia
has enabled the discovery of stellar ensembles
that are more diffuse, but which nonetheless share a common age based
on e.g.~isochrone, rotation, and lithium dating
\cite[e.g.][]{2019AJ....158..122K,Bouma_2022b,Barber_2022}.

@@

The point of Figure~\ref{fig:agescalecompone} is hopefully clear:
until this study, the ages of the youngest Kepler stars and planets were not 
well-measured.
While previous studies reported accurate ages for
$\gtrsim$1\,Gyr stars, ages of sub-gigayear stars have historically been
overestimated by 0.3--2\,dex, with severely understated uncertainties.  
For isochrone and kinematic ages, this is because these
methods rely on parameters that do not appreciably change at
$t$$\lesssim$1\,Gyr.  For the rotation-based ages in
\citet{2023ApJ...952..131M} and \citet{2024AJ....167..159L}, the
discrepancy is
caused by details of the calibration methodologies.  The
\texttt{kiauhoku} spin-down model for instance has known
``stretches'' and ``compressions'' in its age scale relative to observed open clusters
\citep[see][Sec. 7.3]{2023ApJ...952..131M}.
The \citet{Mamajek_2008}
calibration used by \citet{Reinhold_2015} appears more accurate than the former
two models, because it 
was directly fit to reproduce the cluster sequences known at the time.  However the
uncertainties from all of these methods appear to be underestimated.  This
is probably because
they do not marginalize over
the range of rotation periods accessible to stars of a fixed
mass and age.

@@

The age distributions of Kepler stars and KOI hosts in Figure~\ref{fig:hist_tgyro} appear
similar.  One diagnostic for whether the two
distributions are drawn from the same underlying distribution is the
Kolmogorov-Smirnov test;  we calculated this statistic in a manner
that accounts for the Poisson uncertainties by performing 1000 random
draws of 70\% of the stars from each sample.  The resulting 
range of $\log_{\rm 10}p$ (2.5$^{\rm th}$ to 97.5$^{\rm th}$ percentile)
spanned -3.9 to -1.2 for the K dwarfs,
and -5.2 to -2.2 for all Kepler stars.  This degree of similarity suggests 
that although weak differences may be present,
the ``common wisdom'' that young transiting planets are hard to detect
due to the variability of their host stars requires some
nuance.  If true, this statement only holds for the very youngest
stars, where we see a marginal deficit in the KOI host age
distribution relative to that of the overall stellar sample.
However at the age resolution of this study, 
this selection bias against sub-gigayear planets seems small.

@@

We can quantify the relative counts of stars as a function of age by
labelling stars between 0-1\,Gyr, 1-2\,Gyr, and 2-3\,Gyr as ``young'',
``intermediate-age'', and ``old''.  A simple counting exercise from
Figure~\ref{fig:hist_tgyro} tells us that there are \ratioobtoybstars\
times as many old stars in the Kepler field as young stars.
Similarly, there are \ratioobtoybplanets\ times as many old planet
hosts as young planet hosts.  Focusing only on the tails of the
distributions (0-0.3\,Gyr and 2.7-3\,Gyr), the implication is that the
formation rate of stars in the Kepler field decreased by a factor of
\ratiosfr$\pm$\uncratiosfr\ over the past three billion years.
If this decrease continues linearly into the future, then
a simple extrapolation would imply
that in $\approx$1.4$\pm$0.3\,Gyr the rate of new stars forming in the
Galaxy will reach zero.
